Improved Cation Binding to Lipid Bilayer with Negatively Charged POPS by Effective Inclusion of Electronic Polarization

### 2019-10-21 Manuscript/manuscript.tex
Update text

Electrostatic interactions between charges in a dielectric medium are attenuated by a factor equal to the
corresponding relative permittivity (dielectric constant). Standard non-polarizable force fields are capable to
account for the dielectric response from the orientation of molecules (low-frequency), but fail to account for the electronic (high frequency) polarization.
In most biologically relevant environments this high frequency dielectric constant is ~2 (for water the value is 1.78).
This means that for a non-polarizable force field that would strictly account only for the low-frequency dielectric response
interactions between charges in aqueous environments should be further attenuated by a factor of 1.78, which is mathematically
equivalent by scaling the charges by a factor of ~0.75 \cite{leontyev09,leontyev14}. 
This electronic continuum correction (ECC), which can be view as a mean-field continuum way for accounting for electronic polarization,
has been shown to significantly improve the description of interactions between aqueous ions and biologically relevant charged groups \cite{pluharova13,kohagen14,kohagen14b,melcr18,xie16,kroutil17,bruce18}
%[dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp5005693, dx.doi.org/10.1021/jz502099g, DOI: 10.1021/acs.jpcb.7b12510, https://doi.org/10.1021/acs.jpcb.5b10755, DOI:10.1007/s00894-017-3490-x, https://doi.org/10.1063/1.5017101].
This correction is particularly important for high charge density species where it helps to remove artifacts such as excessive ion pairing in water [dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp310719g]. One should, however, bear in mind that existing non-polarizable force fields have been parameterized also using experimental data, which means that they may to a certain extent include implicitly electronic polaization effects. Therefore, in such cases a charge scaling factor between 0.75 and 1 should carefully be adjusted such that electronic polarization effect are not overaccounted.    
In addition to the charge, the Lennard-Jones parameter~$\sigma$ of the ions needed to be modified 
to improve the description of hydration properties of ions with respect to scattering data in these studies 
by scaling with a factor $ 0.72 < f_\sigma \leq 1$.
The reason for the small readjustment of the ionic radii is that the original force fields were often developed such that the first peak on the experimental ion-water oxygen radial distribution function was reproduced. 
Upon charge scaling this peak typically moved to slightly larger distances and the agreement with experiment was then restored by a small decrease of the Lennard-Jones~$\sigma$ parameter.
%More detailed description of ECC and its applications are contained in references \citenum{melcr18, martinek17, Pluharova2014, leontyev09, leontyev11, leontyev14}.

@@

\subsection{Binding of counterions to POPS and POPC and interactions between their headgroups}
Binding affinity of monovalent ions to lipid bilayers depends strongly on
force field parameters in simulations \cite{catte16,NMRlipidsIV}.
The cation binding affinities to lipid bilayers in simulations have been previously evaluated
against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:delta_ordPar_monoval_PCPS} for the labeling) \cite{catte16,melcr18,NMRlipidsIV}.
According to the ''electrometer concept'' \citep{seelig87}, these order parameters
decrease proportionally to the amount of bound positive charge because the headgroup dipole tilts more
parallel to the mebrane normal.
Although the response of PS lipid headgroup order parameters to the bound charge is also systematic,
it is less well undestood and the addition of cations may cause phase transitions in negatively charged
bilayers \cite{feigenson86,mattai89,roux91,roux90}.
Therefore, the cation binding affinity to bilayers with PS lipids has been evaluated by measuring the decrease
of PC headgroup order parameters from POPC:POPS (5:1) mixtures upon addition of salt concentration,
which correlate with the amount of bound cations to lipid bilayer according to the
electrometer concept~\cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV}.
Because the evaluation of monovalent ion binding affinity to bilayers with
PS lipids is complicated by the lack of ion-free state (counterions are always present with negatively charged lipids),
we evaluate the counterion binding affinity to POPC:POPS mixtures also by monitoring
the response of POPC headgroup order parameters to the increasing molar fraction of PS lipids~\cite{NMRlipidsIV} 
(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).

>>>>

Electrostatic interactions between charges in a dielectric medium are attenuated by a factor equal to the
corresponding relative permittivity (dielectric constant). Standard non-polarizable force fields are capable to
account for the dielectric response from the orientation of molecules (low-frequency), but fail to account for the electronic (high frequency) polarization.
In most biologically relevant environments this high frequency dielectric constant is ~2 (for water the value is 1.78).
This means that for a non-polarizable force field that would strictly account only for the low-frequency dielectric response
interactions between charges in aqueous environments should be further attenuated by a factor of 1.78, which is mathematically
equivalent by scaling the charges by a factor of ~0.75 \cite{leontyev09,leontyev14}. 
This electronic continuum correction (ECC), which can be view as a mean-field continuum way for accounting for electronic polarization,
has been shown to significantly improve the description of interactions between aqueous ions and biologically relevant charged groups \cite{pluharova13,kohagen14,kohagen14b,melcr18,xie16,kroutil17,bruce18}
%[dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp5005693, dx.doi.org/10.1021/jz502099g, DOI: 10.1021/acs.jpcb.7b12510, https://doi.org/10.1021/acs.jpcb.5b10755, DOI:10.1007/s00894-017-3490-x, https://doi.org/10.1063/1.5017101].
This correction is particularly important for high charge density species where it helps to remove artifacts such as excessive ion pairing in water~\cite{pluharova13,vazdar13}.
%[dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp310719g].
One should, however, bear in mind that existing non-polarizable force fields have been parameterized also using experimental data, which means that they may to a certain extent include implicitly electronic polaization effects. Therefore, in such cases a charge scaling factor between 0.75 and 1 should carefully be adjusted such that electronic polarization effect are not overaccounted.    
In addition to the charge, the Lennard-Jones parameter~$\sigma$ of the ions needed to be modified 
to improve the description of hydration properties of ions with respect to scattering data in these studies 
by scaling with a factor $ 0.72 < f_\sigma \leq 1$.
The reason for the small readjustment of the ionic radii is that the original force fields were often developed such that the first peak on the experimental ion-water oxygen radial distribution function was reproduced. 
Upon charge scaling this peak typically moved to slightly larger distances and the agreement with experiment was then restored by a small decrease of the Lennard-Jones~$\sigma$ parameter.
%More detailed description of ECC and its applications are contained in references \citenum{melcr18, martinek17, Pluharova2014, leontyev09, leontyev11, leontyev14}.

@@

\subsection{Binding of counterions to POPS and POPC and interactions between their headgroups}
Binding affinity of monovalent ions to lipid bilayers depends strongly on
force field parameters in simulations \cite{catte16,NMRlipidsIV}.
The cation binding affinities to lipid bilayers in simulations have been previously evaluated
against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:delta_ordPar_monoval_PCPS} for the labeling) \cite{catte16,melcr18,NMRlipidsIV}.
According to the ''electrometer concept'' \citep{seelig87}, these order parameters
decrease proportionally to the amount of bound positive charge because the headgroup dipole tilts more
parallel to the membrane normal.
Although the response of PS lipid headgroup order parameters to the bound charge is also systematic,
it is less well undestood and the addition of cations may cause phase transitions in negatively charged
bilayers \cite{feigenson86,mattai89,roux91,roux90}.
Therefore, the cation binding affinity to bilayers with PS lipids has been evaluated by measuring the decrease
of PC headgroup order parameters from POPC:POPS (5:1) mixtures upon addition of salt concentration,
which correlate with the amount of bound cations to lipid bilayer according to the
electrometer concept~\cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV}.
Because the evaluation of monovalent ion binding affinity to bilayers with
PS lipids is complicated by the lack of ion-free state (counterions are always present with negatively charged lipids),
we evaluate the counterion binding affinity to POPC:POPS mixtures also by monitoring
the response of POPC headgroup order parameters to the increasing molar fraction of PS lipids~\cite{NMRlipidsIV} 
(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
